Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-03-10 draft.tex

As a statistical distribution acting on timescales from tens of Myr to several Gyr, the DTD is non-trivial to measure and various techniques have been developed to infer it indirectly. One such approach is to measure global properties of SN Ia host galaxies. Simple examples of such analyses are comparisons of the rates of SNe Ia in hosts that have been split by some observational property, such as morphology or colour, and have shown that SN Ia rate per unit stellar mass (SNuM) is significantly larger in late-type spiral galaxies as well as in bluer galaxies \citep[e.g.][]{Mannucci2005}. These observations were interpreted as showing that SNe Ia are strongly influenced by recent or ongoing star-formation, and thus that the majority of SNe Ia explode after short delay times. Conversely, the SNuM in E/S0 and red galaxies is non-negligible, suggesting that there is a secondary, much older component to the DTD \citep{Sullivan2006,Li2011a,Smith2012}. The DTD was thus approximated to first order by a two-component ("$A + B$") model, where $A$ and $B$ are the normalisations of the DTD for "prompt" (i.e. proportional to instantaneous SFR) and "tardy" (i.e. proportional to overall stellar mass) SNe, respectively.

@@

\citet{Childress2014} showed that the $A+B$ parameterisation arises as a direct consequence of the combination of a power-law DTD with the average SFHs of galaxies. The recent advances of integral field spectroscopy (IFS) allows for an extension of the SFHR method, by reconstructing the SFH for hundreds or thousands of local regions of each SN Ia host galaxy. Using this method, \citet{Castrillo2020} find a power-law slope of $-1.1\pm0.3$, while \citet{Chen2021} find $-1.4\pm0.3$.

@@

\label{sec:data}
\subsection{Dark Energy Survey supernova programme \label{subsec:des}}
The DES Supernova Programme (DES-SN) is a transient survey based on five six-month seasons of observations with the Dark Energy Camera (DECam; \citealt{Flaugher2015}), designed primarily to measure the light curves of SNe Ia for use as cosmological distance indicators. Transients were detected and processed using a difference imaging program \citep{Kessler2015} and vetted using a machine learning (ML) algorithm \citep{Goldstein2015}. Spectral follow-up of live SN candidates was performed on a suite of large optical telescopes \citet{Smith2020a}, leading to the initial publication of a measurement of cosmological parameters using 207 spectroscopically confirmed SNe Ia (\citealt{DESCollaboration2018a}, and references therein). 
\subsection{Supernovae \label{subsec:host_sample}}
\subsubsection{Photometric classification and quality cuts \label{subsubsec:sn_classify}}
With over 30,000 transients passing ML selection it was not possible to obtain spectral follow up of every object. Instead we make use of photometric classification in the form of the recurrent neural network classifier \texttt{superNNova} (SNN; \citealt{Moller2019}), mirroring the approach of \citet{Scolnic2020}. We train SNN using a large suite of simulated multi-band SN light curves. \textbf{Waiting on final sample and a few words from Maria/Anais to complete this bit.} SNe Ia are simulated based on the \texttt{SALT2} model \citep{Guy2007} trained on the Joint Lightcurve Analysis (JLA) data set \citep{Betoule2014}. CCSNe are generated from the templates of \citet{Vincenzi2019}. We select SNe with a threshold probability of $P(\mathrm{Ia})\geq0.5$, although the our analysis is not sensitive to this choice, since the vast majority of SNe receive classifications close to 1 or 0. The photometrically classified SNe Ia are then passed through the \texttt{SALT2} framework light curve fitter and are subject to quality cuts on the light curve width (stretch; $x_1$) and colour ($c$) parameters in an identical manner to \citet{Vincenzi2020}: $-3 \leq x_1 \leq 3$, $-0.3 \leq c \leq 0.3$. This selection helps to reduce the potential contamination from core-collapse SNe (CCSNe) and outlying thermonuclear SNe. We impose further cuts on the quality of those measured parameters, which are: the uncertainty on $x_1$: $\sigma_{x_1} <1$; the uncertainty on the date at which the light curve peaked: $\sigma_{t_{\mathrm{peak}}} <2$; the SALT2 fit probability $>0.001$.

\subsubsection{Host galaxy selection \label{subsubsec:sn_hosts}}
Host galaxies for all transients in DES-SN are retrieved from the DES-SN Deep catalogue, which is derived from the $g, r, i, z$-band coadded images presented in \citet{Wiseman2020}. SN are associated to hosts using the directional light radius method \citep[e.g.][]{Sullivan2006,Gupta2016}, and spectroscopic redshifts for host galaxies come from the OzDES Global Redshift Catalog\footnote{\url{https://docs.datacentral.org.au/ozdes/overview/dr2/}} \citep{Yuan2015,Childress2017,Lidman2020}, which comprises galaxies targetted as DES-SN hosts with the OzDES programme as well as redshifts from legacy catalogues in the DES-SN fields.

@@

To compare the properties of SN hosts to the general galaxy population, we select a sample of field galaxies. In order to maintain consistency between the SN host and field galaxy sample selections, we obtain a field galaxy sample from the same catalogue as that from which the SNe are matched to obtain host galaxies. The steps relevant to this analysis are outlined in the following sections.

@@

Since the vast majority of objects detected in the SN Deep catalogue do not have a spectroscopic redshift measurement, we rely on photometric redshifts (photo-$z$s).
Photo-$z$s are taken, where available, from the Y3A2\_DEEP catalogue of \citet{Hartley2020}, and have been estimated using the \texttt{eaZy} code \citep{Brammer2008}. \citet{Hartley2020} use deep DES optical photometry alongside DECam $u$-band data and near-infrared (NIR) $J$, $H$, and $K$-band imaging from the ultraVISTA and VIDEO surveys. The deep optical photometry in Y3A2\_DEEP was stacked using a different technique to that of SN Deep, but the resulting photometry is consistent within $1\sigma$ uncertainties \citep{Wiseman2020}. Photo-$z$ accuracy from \citet{Hartley2020} at $17 < i<24$ is estimated around 0.03 as quantified by the Normalised Absolute Median Deviation (NMAD). This accuracy is degraded to 0.07 at $17 < i<26$, a range that includes all galaxies in our sample, but is more than adequate for the purposes of this work, where all field galaxies will be grouped into stellar mass bins of 0.5 dex.

@@

\begin{enumerate}
    \item objects must be detected and have a Kron magnitude measurement in all four DES optical bands

    \item field galaxies are limited to unmasked region of the SN-X3 chip \citep{Hartley2020}

    \item objects must not be within a given number (20 in $x$, 50 in $y$) of pixels of the CCD edge, as the co-addition of slightly misaligned images introduced a region of significant noise in this part of the detector

    \item to minimise stellar contamination, objects must have a value of 0.95 in the star/galaxy (S/G) separation \texttt{CLASS\_STAR} provided by \texttt{Source Extractor} \citep{Bertin1996}

    \item field galaxies must be covered by $ugrizJHK$ photometry with either detections or upper limits

    \item objects must have a well-defined spectroscopic or photometric redshift measurement

    \item objects must be detected at signal-to-noise ratio (SNR) greater than 3 in the $r$ band

    \item galaxies must be brighter than $m_r \leq 24.5$, as this is the magnitude of the faintest SN host with a spectroscopic redshift.

@@

We estimate global galaxy properties for both the SN host and field galaxies that pass the quality cuts by fitting the photometry with stellar population templates in a method outlined by \citet{Sullivan2006} and consistent to that used in previous DES-SN analyses \citep{kelsey2021,Smith2020,Wiseman2020}. For SN hosts the redshift is fixed at the spectroscopically determined value, while for field galaxies we fix it at either the spectroscopic value if one exists in the OzDES GRC, or more commonly the photometrically derived value. We use the stellar population templates of \citet{Bruzual2003} and adopt a \citet{Chabrier2003} initial mass function (IMF). The fitting procedure returns a best fitting template and corresponding stellar mass ($M_*$). Upper and lower bounds on stellar mass estimates are taken as the largest and smallest values of a given parameter for which the reduced $\chi^2$ parameter corresponds to $\chi^2_{\mathrm{red, min}} +2.3$. To check that the choice of templates do not bias the result, we also fit the galaxies using the P\'EGASE.2 templates \citep{Fioc1997,LeBorgne2002} and a \citet{Kroupa2001} IMF, and find results consistent within measurement uncertainties. This is consistent with the findings of \citet{Smith2020}.

The results of our SED fitting are shown in Fig. \ref{fig:mass_hist_hosts_field}. The upper panel, with both the SN hosts and field galaxies displayed, shows the vastly different distributions of the two samples. SN hosts are preferentially high mass galaxies, whereas the field galaxy distributions increases down to lower masses, peaking around $10^{8.5}~\mathrm{M}_{\odot}$. In the bottom panel, the same field galaxy stellar mass distribution is shown on a log scale, along with the stellar mass function (SMF) from the ZFOURGE survey \citep{Tomczak2014}, whose stellar masses are derived with the same photo-$z$ code, IMF and template library as used here. Above $10^{8.5}~\mathrm{M}_{\odot}$, the DES field galaxy distribution closely follows the ZFOURGE SMF for $0.5<z<0.75$, indicating that the DES sample is representative of field galaxies in this redshift range.%To estimate uncertainties on the physical properties, we resample the photometry from a Gaussian distribution formed by the flux measurements and their uncertainties and rerun the SED fit. This procedure is repeated 1000 times. 
%%%% INCOMPLETENESS %%%%

\section{Incompleteness corrections}
\label{sec:incompleteness}
The simple ratio of the number of SN hosts and field galaxies presented at the end of the previous section provides a first approximation of the SN rate per galaxy, and the addition of factor equal to the survey duration normalises the rate to per year. However both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources incompleteness in the data.

@@

To correct for these incompleteness, we follow a similar method to that used in the PTF rates analyses of \citet{Frohmaier2019} and \citet{Frohmaier2020}. We run a suite of simulations using an identical method to that described in detail in \citet{Vincenzi2020}. We simulate SNe using the SuperNova ANAlysis software (\texttt{SNANA} \citealt{Kessler2009a}) integrated into the \texttt{pippin} framework \citep{Hinton2020}. We simulate \textbf{size of final sim} SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$. The SNe are generated and trained in the same way as for the training of SNN. SALT2 parameters $x_1$ (stretch) and $c$ (colour) are drawn randomly from the intrinsic asymmetric Gaussian distributions described in \citet{Scolnic2016} and "blurred" following the intrinsic scatter model of \citet{Guy2010}, while redshifts are drawn following the volumetric rate evolution of \citet{Frohmaier2019}. The SNe are simulated with explosion epochs $t_0$ uniformly distributed between two months before DES-SN began and two months after it finished. We run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on a combination of sky location, explosion epoch, redshift, stretch, and colour. The fraction of recovered SNe (the efficiency) is thus described by a 5-dimensional surface. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, with stretch $x_1$ and colour $c$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i},x_1,c) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i},x_1,c\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i},x_{1,i},c_i\right)}\right)\,.
\end{equation}

@@

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band \citep{Lidman2020}. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citet{Vincenzi2020} who provide the spectroscopic redshift efficiency as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. So that we reduce any bias towards SNe in bright hosts that are easier to obtain spectroscopic redshifts for, we assign each SN in the sample described in Section \ref{subsec:host_sample} a weight equal to the inverse of the spectroscopic efficiency for its host.

@@

\subsubsection{Apparent magnitude limits \label{subsubsec:mag_lims}}
As we use photometric, rather than spectroscopic, redshifts for the field galaxies, they do not suffer from spectroscopic incompleteness as the SN hosts do. Instead, the inclusion of any given galaxy in the survey area in the sample is determined simply by whether it is detected above a prescribed threshold in signal-to-noise ratio, i.e. the sample is magnitude limited modulo the cuts described in Section \ref{subsec:cuts}. To determine the apparent magnitude limit in each of the optical bands we employ the method of \citet{Johnston2007}, \citet{Teodoro2010}, and \citet{Johnston2012} (hereafter Completeness I, II, III respectively). Test statistics $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ are computed based on absolute magnitude and distance modulus by calculating the rank of each galaxy's absolute magnitude (distance modulus) when compared to all other galaxies in a survey within a certain slice of distance modulus (absolute magnitude). For a galaxy of apparent magnitude $m < m_{\mathrm{lim, trial}}$ observed in a survey complete to magnitude $m_{\mathrm{lim, true}}$ where  $m_{\mathrm{lim, trial}} \leq m_{\mathrm{lim, true}}$, the expectation value of the rank is 0.5. However, if a trial limiting magnitude $m_{\mathrm{lim, trial}} \geq m_{\mathrm{lim, true}}$, there will be a lack of observed faint objects, such that the expectation value of the rank drops. Completeness I, II, and III show that for surveys with sharp, well defined magnitude limits the test statistics $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ drop sharply. The DES deep stacks present a more challenging case. They cover a wide area and are compiled from tens of separate CCDs, each with its own detection efficiency. Moreover, galaxy redshifts have been estimated using photometry only, and $K$-corrections performed using an imperfect template fit, potentially leading to large systematic and statistical fluctuations in absolute magnitudes and distance moduli. To reduce the inhomogeneity while still including  large sample of objects, we split the sample into the deep (X3) and shallow (E2) fields and calculate limiting magnitudes separately. Fig. \ref{fig:completeness} shows the values of $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$. The values of the test statistics increase with $m_{\mathrm{lim, trial}}$ until a peak, before decreasing to stable values at magnitudes far beyond the limit of the survey. The shape at brighter $m_{\mathrm{lim, trial}}$ is likely caused by incompleteness at the bright end: due to the small sky area, we simply do not probe enough volume to sample the bright end of the galaxy luminosity function well enough for the statistics to be robust. To approximate an efficiency function, we fit the peak of the statistics with a polynomial function, and then normalise by the maximum and minimum values. We then interpolate between the peak and the faint-magnitude floor to find the 50\% completeness limit at the point where the normalised value of the test statistic is 0.5. These values are consistent with the value at which 50\% of the true sources are detected but have the advantage of being derived from the data without a simulation that is based on assumptions and thus susceptible to bias.

@@

\section{The per-galaxy rate of type Ia supernovae}
\label{sec:rates}
The rate of SNe per galaxy ($R_{\mathrm{G}}$) per year in a transient survey can be approximated by the equation:
\begin{equation}
    R_{\mathrm{G}} = \frac{N_{\mathrm{SN}}}{N_{\mathrm{G}}} \frac{V_{\mathrm{G}}}{V_{\mathrm{SN}}}\frac{1}{T_{\mathrm{SN  survey}}}\,,
\label{eq:rate1}
\end{equation}
where $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ are the respective number of SNe and galaxies detected, $V_{\mathrm{SN}}$ and $V_{\mathrm{G}}$ the volumes from which the SNe and G samples were taken, and $T$ the duration of the SN survey in years.

@@

\begin{equation}
    N_{\mathrm{SN,intrinsic}} = \sum_i^{n_{\mathrm{SN}}}\eta_{\mathrm{SN,}i}\,,
    \label{eq:corr_SN}
\end{equation}
and
\begin{equation}
    N_{\mathrm{G,intrinsic}} = \sum_j^{n_{\mathrm{G}}} \eta_{\mathrm{G,}j} 
    \label{eq:corr_G}
\end{equation}
for each SN $i$ and galaxy $j$.

@@

In Section \ref{sec:rates} and Fig. \ref{fig:rate_raw} we showed that the SN Ia rate per galaxy as a function of stellar mass is approximated by, but not perfectly described by, a power-law (and thus a linear relationship in log space). In this section, we introduce a physical model in order to better fit the observed rate vs stellar mass relationship.

@@

As per \citet{Childress2014} we plant seed galaxies with initial masses of $10^6 \mathrm{M}_{\odot}$ at intervals of 50 Myr for look-back times $1 \leq t_f \leq 10$ Gyr, and 25 Myr for look-back times $10 \leq t_f \leq 13$ Gyr, and evaluate the model with a time step of 0.5 Myr.

>>>>

As a statistical distribution acting on timescales from tens of Myr to several Gyr after an epoch of star formation, the DTD is non-trivial to measure and various techniques have been developed to infer it indirectly. One such approach is to measure global properties of SN Ia host galaxies. Simple examples of such analyses are comparisons of the rates of SNe Ia in hosts that have been split by some observational property, such as morphology or colour, and have shown that SN Ia rate per unit stellar mass (SNuM) is significantly larger in late-type spiral galaxies as well as in bluer galaxies \citep[e.g.][]{Mannucci2005}. These observations were interpreted as showing that SNe Ia are strongly influenced by recent or ongoing star-formation, and thus that the majority of SNe Ia explode after short delay times. On the other hand, the SNuM in E/S0 and red galaxies is also non-negligible, suggesting that there is a secondary, much older component to the DTD \citep{Sullivan2006,Li2011a,Smith2012}. The DTD was thus approximated to first order by a two-component ("$A + B$") model, where $A$ and $B$ are the normalisations of the DTD for "prompt" (i.e. proportional to instantaneous SFR) and "tardy" (i.e. proportional to overall stellar mass) SNe, respectively.

@@

\citet{Childress2014} showed that the $A+B$ approximation arises as a direct consequence of the combination of a power-law DTD with the average SFHs of galaxies -- the prompt component proportional to the amount of on-going star-formation at the epoch of observation, and the tardy component caused by the fact that massive galaxies experienced high SFRs several Gyrs ago. The recent advances of integral field spectroscopy (IFS) allows for an extension of the SFHR method, by reconstructing the SFH for hundreds or thousands of local regions of each SN Ia host galaxy. Using this method, \citet{Castrillo2020} find a power-law slope of $-1.1\pm0.3$, while \citet{Chen2021} find $-1.4\pm0.3$.

@@

\label{sec:data}
\subsection{Dark Energy Survey supernova programme \label{subsec:des}}
The Dark Energy Survey Supernova Programme (DES-SN) is a transient survey based on five six-month seasons of observations with the Dark Energy Camera (DECam; \citealt{Flaugher2015}), designed primarily to measure the light curves of SNe Ia for use as cosmological distance indicators. Transients were detected and processed using a difference imaging program \citep{Kessler2015} and vetted using a machine learning (ML) algorithm \citep{Goldstein2015}. Spectral follow-up of live SN candidates was performed on a suite of large optical telescopes \citep{Smith2020a}, leading to the initial publication of a measurement of cosmological parameters using 207 spectroscopically confirmed SNe Ia (\citealt{DESCollaboration2018a}, and references therein). 
\subsection{Supernovae \label{subsec:host_sample}}
\subsubsection{Photometric classification and quality cuts \label{subsubsec:sn_classify}}
With over 30,000 transients passing ML selection it was not possible to obtain spectral follow up of every object. Instead we make use of photometric classification in the form of the recurrent neural network classifier \texttt{superNNova} (SNN; \citealt{Moller2019}), mirroring the approach of \citet{Scolnic2020}. We train SNN using a large suite of simulated multi-band SN light curves. \textbf{Waiting on final sample and a few words from Maria/Anais to complete this bit.} SNe Ia are simulated based on the \texttt{SALT2} model \citep{Guy2007} trained on the Joint Lightcurve Analysis (JLA) data set \citep{Betoule2014}. CCSNe are generated from the templates of \citet{Vincenzi2019}. We select SNe with a threshold probability of $P(\mathrm{Ia})\geq0.5$, although our analysis is not sensitive to this choice, since the vast majority of SNe receive classifications close to 1 or 0. The photometrically classified SNe Ia are then passed through the \texttt{SALT2} framework light curve fitter and are subject to quality cuts on the light curve width (stretch; $x_1$) and colour ($c$) parameters in an identical manner to \citet{Vincenzi2020}: $-3 \leq x_1 \leq 3$, $-0.3 \leq c \leq 0.3$. This selection helps to reduce the potential contamination from core-collapse SNe (CCSNe) and outlying thermonuclear SNe. We impose further cuts on the quality of those measured parameters, which are: the uncertainty on $x_1$: $\sigma_{x_1} <1$; the uncertainty on the date at which the light curve peaked: $\sigma_{t_{\mathrm{peak}}} <2$; the SALT2 fit probability $>0.001$.

\subsubsection{Host galaxy selection \label{subsubsec:sn_hosts}}
Host galaxies for all transients in DES-SN are retrieved from the DES-SN Deep catalogue, a galaxy catalogue from the $g, r, i, z$-band coadded images of the ten DES-SN fields which was presented in \citet{Wiseman2020}. SNe are associated to galaxies using the directional light radius (DLR) method \citep[e.g.][]{Sullivan2006,Gupta2016}. As is standard across DES-SN analyses, we require a SN-host separation to DLR ratio ($d_{\mathrm{DLR}}$) less than 4 in order to classify a galaxy as a host. In the case where multiple galaxies lie within $d_{\mathrm{DLR}}<4$, the object with the lower or lowest value is taken to be the host. Spectroscopic redshifts for host galaxies come from the OzDES Global Redshift Catalog (GRC; \footnote{\url{https://docs.datacentral.org.au/ozdes/overview/dr2/}} \citealt{Yuan2015,Childress2017,Lidman2020}), which comprises galaxies targetted as DES-SN hosts with the OzDES programme as well as redshifts from legacy catalogues in the DES-SN fields.

@@

To calculate the SN Ia rate per galaxy as a function of stellar mass we require a representative sample of the global galaxy population, which we call our field galaxy sample. In order to maintain consistency between the SN host and field galaxy sample selections, we obtain a field galaxy sample from the same catalogue as that from which the SNe are matched to obtain host galaxies (SN Deep; \citealt{Wiseman2020}). The steps relevant to this analysis are outlined in the following sections.

@@

Since the vast majority of objects detected in the SN Deep catalogue do not have a spectroscopic redshift measurement, we rely on photometric redshifts (photo-$z$s).
Photo-$z$s are taken, from the DES Y3A2\_DEEP catalogue of \citet{Hartley2020}. Y3A2\_DEEP makes use of the same deep DES optical photometry that was used in SN Deep for a subset of the fields (SN-X3, SN-C3 and SN-E2), but adds DECam $u$-band data and near-infrared (NIR) $J$, $H$, and $K$-band imaging from the ultraVISTA and VIDEO surveys. The deep optical photometry in Y3A2\_DEEP was stacked using a different technique to that of SN Deep, but the resulting photometry is consistent within $1\sigma$ uncertainties \citep{Wiseman2020}. \citet{Hartley2020} estimated photo-$z$'s using the \texttt{eaZy} code \citep{Brammer2008}. Photo-$z$ accuracy from \citet{Hartley2020} at $17 < i<24$ is estimated around 0.03 as quantified by the Normalised Absolute Median Deviation (NMAD). This accuracy is degraded to 0.07 at $17 < i<26$, a range that includes all galaxies in our sample, but is more than adequate for the purposes of this work, where all field galaxies will be grouped into stellar mass bins of $\log \left(\Delta M_*\right) = 0.25$.

@@

\begin{enumerate}
    \item objects must be detected and have a Kron magnitude measurement in all four DES optical bands;

    \item field galaxies are limited to unmasked region of the SN-X3 chip \citep{Hartley2020};

    \item objects must not be within a given number (20 in $x$, 50 in $y$) of pixels of the CCD edge, as the co-addition of slightly misaligned images introduced a region of significant noise in this part of the detector;

    \item to minimise stellar contamination, objects must have a value of 0.95 in the star/galaxy (S/G) separation \texttt{CLASS\_STAR} provided by \texttt{Source Extractor} \citep{Bertin1996};

    \item field galaxies must lie within the area covered by $ugrizJHK$ photometry, with a photometric flux measurement or upper limit present in all bands;

    \item objects must have a spectroscopic redshift measurement or a converged photometric redshift estimate;

    \item objects must be detected at signal-to-noise ratio (SNR) greater than 3 in the $r$ band;

    \item galaxies must be brighter than $m_r \leq 24.5$, as this is the magnitude of the faintest SN host with a spectroscopic redshift;

@@

We estimate global galaxy properties for both the SN host and field galaxies that pass the quality cuts by fitting the photometry with stellar population templates in a method outlined by \citet{Sullivan2006} and consistent to that used in previous DES-SN analyses \citep{kelsey2021,Smith2020,Wiseman2020}. For SN hosts the redshift is fixed at the spectroscopically determined value, while for field galaxies we fix it at either the spectroscopic value if one exists in the OzDES GRC, or more commonly the photometrically derived value. We use the stellar population templates of \citet{Bruzual2003} and adopt a \citet{Chabrier2003} initial mass function (IMF). The fitting procedure returns a best fitting template and corresponding stellar mass ($M_*$). Upper and lower bounds on stellar mass estimates are taken as the extrema in the model parameters that are consistent with the data (given the photometric uncertainties) according to their $\chi^2$ statistic, as per \citet{Sullivan2006}. To check that the choice of templates does not bias the result, we also fit the galaxies using the P\'EGASE.2 templates \citep{Fioc1997,LeBorgne2002} and a \citet{Kroupa2001} IMF, and find results consistent within measurement uncertainties. This is consistent with the findings of \citet{Smith2020}.

The results of our SED fitting are shown in Fig. \ref{fig:mass_hist_hosts_field}. The upper panel, with both the SN hosts and field galaxies displayed, shows the vastly different distributions of the two samples. SN hosts are preferentially high mass galaxies, whereas the field galaxy distribution increases down to lower masses, peaking around $10^{8.5}~\mathrm{M}_{\odot}$. In the bottom panel, the same field galaxy stellar mass distribution is shown on a log scale, along with the stellar mass function (SMF) from the ZFOURGE survey \citep{Tomczak2014}, whose stellar masses are derived with the same photo-$z$ code, IMF and template library as used here. Above $10^{8.5}~\mathrm{M}_{\odot}$, the DES field galaxy distribution closely follows the ZFOURGE SMF for $0.5<z<0.75$, indicating that the DES sample is representative of field galaxies in this redshift range.%To estimate uncertainties on the physical properties, we resample the photometry from a Gaussian distribution formed by the flux measurements and their uncertainties and rerun the SED fit. This procedure is repeated 1000 times. 
%%%% INCOMPLETENESS %%%%

\section{Incompleteness corrections}
\label{sec:incompleteness}
The simple ratio of the number of SN hosts and field galaxies presented at the end of the previous section provides a first approximation of the SN rate per galaxy, and the addition of a factor equal to the survey duration normalises the rate to per year. However, both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources incompleteness in the data.

@@

To correct for these incompleteness, we follow a similar method to that used in the Palomar Transient Factory rates analyses of \citet{Frohmaier2019} and \citet{Frohmaier2020}. We run a suite of simulations using an identical method to that described in detail in \citet{Vincenzi2020}. We simulate SNe using the SuperNova ANAlysis software (\texttt{SNANA} \citealt{Kessler2009a}) integrated into the \texttt{pippin} framework \citep{Hinton2020}. We simulate \textbf{size of final sim} SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$. The SNe are generated and trained in the same way as for the training of SNN. SALT2 parameters $x_1$ (stretch) and $c$ (colour) are drawn randomly from the intrinsic asymmetric Gaussian distributions described in \citet{Scolnic2016} and "blurred" following the intrinsic scatter model of \citet{Guy2010}, while redshifts are drawn following the volumetric rate evolution of \citet{Frohmaier2019}. The SNe are simulated with explosion epochs $t_0$ uniformly distributed between two months before DES-SN began and two months after it finished in order to account for all SNe that could have been observed by DES-SN. We run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on a combination of sky location, explosion epoch, redshift, stretch, and colour. The fraction of recovered SNe (the efficiency) is thus described by a 5-dimensional surface. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, with stretch $x_1$ and colour $c$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i},x_1,c) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i},x_1,c\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i},x_{1,i},c_i\right)}\right)\,.
\end{equation}

@@

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band \citep{Lidman2020}. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citet{Vincenzi2020} who provide the spectroscopic redshift efficiency as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. So that we reduce any bias towards SNe in bright hosts that are easier to obtain spectroscopic redshifts for, the host galaxy spectroscopic redshift efficiency is multiplied by that derived from the SN detection efficiency (Section \ref{subsec:incompleteness_SNe}) to arrive at the final SN efficiency $\eta_{\mathrm{SN},i}$.

@@

\subsubsection{Apparent magnitude limits \label{subsubsec:mag_lims}}
As we use photometric, rather than spectroscopic, redshifts for the field galaxies, they do not suffer from spectroscopic incompleteness as the SN hosts do. Instead, the inclusion of any given galaxy in the survey area in the sample is determined simply by whether it is detected above a prescribed threshold in signal-to-noise ratio, i.e. the sample is magnitude limited modulo the cuts described in Section \ref{subsec:cuts}. To determine the apparent magnitude limit in each of the optical bands we employ the method of \citet{Johnston2007}, \citet{Teodoro2010}, and \citet{Johnston2012} (hereafter Completeness I, II, III respectively). We use the test statistics $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ from those studies, which are computed based on absolute magnitude (distance modulus). Each galaxy is ranked by its absolute magnitude (distance modulus) when compared to all other galaxies in a survey within a certain slice of distance modulus (absolute magnitude). For a galaxy of apparent magnitude $m < m_{\mathrm{lim, trial}}$ observed in a survey complete to magnitude $m_{\mathrm{lim, true}}$ where  $m_{\mathrm{lim, trial}} \leq m_{\mathrm{lim, true}}$, the expectation value of the rank is 0.5. However, if a trial limiting magnitude $m_{\mathrm{lim, trial}} \geq m_{\mathrm{lim, true}}$, there will be a lack of observed faint objects, such that the expectation value of the rank drops.

Completeness I, II, and III show that for surveys with sharp, well defined magnitude limits the test statistics $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ drop sharply. The DES deep stacks present a more challenging case. They cover a wide area and are compiled from tens of separate CCDs, each with its own detection efficiency. Moreover, galaxy redshifts have been estimated using photometry only, and $K$-corrections performed using an imperfect template fit, potentially leading to large systematic and statistical fluctuations in absolute magnitudes and distance moduli. To reduce the inhomogeneity while still including  large sample of objects, we limit our sample to include only data from one deep (SN-X3) field. Fig. \ref{fig:completeness} shows the values of $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$. The values of the test statistics increase with $m_{\mathrm{lim, trial}}$ until a peak, before decreasing to stable values at magnitudes far beyond the limit of the survey. The shape at brighter $m_{\mathrm{lim, trial}}$ is likely caused by incompleteness at the bright end: due to the small sky area, we simply do not probe enough volume to sample the bright end of the galaxy luminosity function well enough for the statistics to be robust. To approximate an efficiency function, we fit the peak of the statistics with a polynomial function, and then normalise by the maximum and minimum values. We then interpolate between the peak and the faint-magnitude floor to find the 50\% completeness limit at the point where the normalised value of the test statistic is 0.5. These values are consistent with the value at which 50\% of the true sources are detected but have the advantage of being derived from the data without a simulation that is based on assumptions and thus susceptible to bias.

@@

\section{The per-galaxy rate of type Ia supernovae}
\label{sec:rates}
The rate of SNe per galaxy ($R_{\mathrm{G}}$) per year in a transient survey can be approximated by the equation:
\begin{equation}
    R_{\mathrm{G}} = \frac{N_{\mathrm{SN}}}{N_{\mathrm{G}}} \frac{V_{\mathrm{G}}}{V_{\mathrm{SN}}}\frac{1}{T}\,,
\label{eq:rate1}
\end{equation}
where $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ are the respective numbers of SNe and galaxies detected, $V_{\mathrm{SN}}$ and $V_{\mathrm{G}}$ the volumes from which the SNe and galaxy samples were taken, and $T$ the duration of the SN survey in years.

@@

\begin{equation}
    N_{\mathrm{SN,intrinsic}} = \sum_i^{n_{\mathrm{SN}}}\eta\left(F,z,t_0,x_1,c,m_{\mathrm{host},r}\right)_{i}\,,
    \label{eq:corr_SN}
\end{equation}
and
\begin{equation}
    N_{\mathrm{G,intrinsic}} = \sum_j^{n_{\mathrm{G}}} \eta\left(V_{\mathrm{max}}\right)_{j} 
    \label{eq:corr_G}
\end{equation}
for each SN $i$ and galaxy $j$.

@@

In Section \ref{sec:rates} and Fig. \ref{fig:rate_raw} we showed that the SN Ia rate per galaxy as a function of stellar mass is approximated by, but not perfectly described by, a power-law. In this section, we introduce a physical model in order to better fit the observed rate vs stellar mass relationship.

@@

Finally, at time $t$, some time $\tau$ since an epoch of star formation, a fraction of stellar mass is lost. We adopt the parametrization used by \citet{Childress2014}, namely that of \citet{Leitner2011} for a Chabrier 2003 IMF:
